Citation network analysis of urban modeling research

### 2019-06-05 paper.tex
Update on Overleaf.

Urban modeling, defined broadly as quantitative modeling and simulation aiming at understanding urban processes, has always been a highly multidisciplinary and interdisciplinary subject. Indeed, several disciplines ranging from Urban Economics~\cite{krugman1998s} to Quantitative Urban Geography~\cite{pumain2017geography} or Urban Planning~\cite{knowles2016sir} provide complementary viewpoints by studying different dimensions of urban phenomena. Despite sometimes converging issues, bridges remain rather rare, such as integrating approaches such as the Science of Cities introduced by \cite{batty2013new}. One crucial question in that context is to what extent these different approaches to modeling urban issues may be complementary, and what may the potentially fertile couplings in terms of producing new knowledge.

% \cite{ramadier2004transdisciplinarity} transdisciplinarity in urban studies
% \cite{batty2019map} the digital twin - need of reflexivity, hence understanding modeling landscape is crucial

@@

% \cite{giatsidis2019rooted} new measures based on subgraph density for citation networks

We construct a large corpus of papers, linked through backward citations.
Our contribution relies on the following points: 
% (i) sensitivity analysis - info to reconstruct discipline with such empirical data ?
% (ii) some methodological insights depending on what we do ?
% (iii) results on urban modeling in themselves <- Maybe insist only on that ?

The rest of the paper is organized as follows. We first describe the construction of the dataset. We then explore it answering different thematic questions; ..

@@

@@

We propose a double entry to build a broad corpus, on complementary but distinct approaches. This will ensure first that the collected data is representative due to the corpus size limitations, but secondly that potential couplings are included. % REFORMULATE

Our corpus is surely not exhaustive to urban modeling as we focused on the aforementioned two main entries, and other requests such as ``urban geosimulation'', ``urban growth model'', ``urban agent based model'' would also be relevant. We expect for example, following \cite{wise2016transportation}, to have a strong overlap with our corpus and a potential corpus related to agent-based modeling.
%urban agent based model, spatial interaction model, spatial interaction agent based model, geosimulation, urban geosimulation, urban cellular automata, urban growth model, urban land use change model, urban land use model

Language statistics (detected using titles). We do not any semantic analysis here and thus do not filter on language.
% possible additional analysis by isolating language subnetworks - checking modularity of language communities ?. Typically: Chinese article are only cited by other Chinese articles. similar but less strong effect with French

% main languages
%    Chinese      Czech      Dutch    English     French   Galician     German Indonesian 
%     24211        589        872     450149       6425        663       8467        880 
%   Italian   Japanese     Korean     Polish Portuguese    Russian    Spanish    Swedish 
%      1641       2457       2080       2284       6348       1981       9234        654 
%   Turkish 
%      1196

@@

@@

We keep the giant clusters %TODO cluster statistics
and the ``core'' of the network, which we define as

%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

@@

Overlap for the different keywords combinations, maybe filter papers that are present in all keyword searches

\subsection*{Endogenous disciplines}

% cluster sizes

>>>>

Urban modeling, defined broadly as quantitative modeling and simulation aiming at understanding urban processes, has always been a highly multidisciplinary and interdisciplinary subject \cite{ramadier2004transdisciplinarity}. Indeed, several disciplines ranging from Urban Economics~\cite{krugman1998s} to Quantitative Urban Geography~\cite{pumain2017geography} or Urban Planning~\cite{knowles2016sir} provide complementary viewpoints by studying different dimensions of urban phenomena. Despite sometimes converging issues, bridges remain rather rare, such as integrating approaches such as the Science of Cities introduced by \cite{batty2013new}. One crucial question in that context is to what extent these different approaches to modeling urban issues may be complementary, and what may the potentially fertile couplings in terms of producing new knowledge. Indeed, a good knowledge of the scientific landscape, or in other terms a certain level of reflexivity, is crucial for urban modeling studies since as \cite{batty2019map} points out, models become digital twins of modelled systems which themselves play a role in the studied system.

@@

%\cite{giatsidis2019rooted} propose  new measures based on subgraph density for citation networks

We construct a large corpus of papers, linked through backward citations.
Our contribution relies on the following points: (i) the open dataset we make available can be of significant interest for researcher interested in urban modeling; (ii) we map a broad part of approaches to urban modeling, and more particularly clarify the relations between microsimulation, land-use transport modeling, and agent-nased modeling.
% (i) sensitivity analysis - info to reconstruct discipline with such empirical data ?
% (ii) some methodological insights depending on what we do ?
% (iii) results on urban modeling in themselves <- Maybe insist only on that ?

The rest of the paper is organized as follows. We first describe the construction of the dataset, i.e. the initial assumptions and data collection process, and the filtering steps to work on a relevant network. We then study its sensitivity to construction parameters, what allows to comment on endogenous disciplines and study the evolution of communities in time. We finally discuss the implication of these results, for example in the context of coupling approaches.

@@

\subsection*{Initial seed}

We propose a multiple keyword entry to build a broad corpus, on complementary but distinct approaches. This will ensure first that the collected data is representative due to the corpus size limitations, but secondly that potential couplings are included. % REFORMULATE

Our corpus is surely not exhaustive to urban modeling as we focused on the aforementioned two main entries, and other requests such as ``urban geosimulation'', ``urban growth model'', ``urban agent based model'' would also be relevant. We expect for example, following \cite{wise2016transportation}, to have a strong overlap with our corpus and a potential corpus related to agent-based modeling. Similarly, other sub-fields of urban modeling such as cellular automata \cite{white1993cellular} or geosimulation \cite{benenson2004geosimulation} are expected to be highly interwined with our corpus from a thematic viewpoint.
%urban agent based model, spatial interaction model, spatial interaction agent based model, geosimulation, urban geosimulation, urban cellular automata, urban growth model, urban land use change model, urban land use model

@@

@@

\subsection*{Network filtering}

Language statistics (detected using titles). We do not include in our analysis any semantic analysis and thus do not filter on language of papers. Note that as we work on citation networks, cross-language citations are a common feature and different languages will not produce isolated clusters. In any case, a language detection with the python package \texttt{polyglot} confirms that most of the corpus is in English (86.6\%), secondary main language being Mandarin (4.7\%), Spanish (1.8\%), German (1.6\%), French (1.2\%) and Portuguese (1.2\%).
% possible additional analysis by isolating language subnetworks - checking modularity of language communities ?. Typically: Chinese article are only cited by other Chinese articles. similar but less strong effect with French

%     Czech    Swedish   Galician      Dutch Indonesian    Turkish   
% 0.1129035  0.1259161  0.1274470  0.1672503  0.1693553  0.2309738  
% Italian    Russian     Korean     Polish 
%0.3155558  0.3832979  0.3987982  0.4384102 
%  Japanese Portuguese     French     German    Spanish  
% 0.4720899  1.2195495  1.2339017  1.6231318  1.7783263 
% Chinese    English 
%4.6506688 86.5524236

@@

We keep the giant connected cluster of the resulting network, which covers 99.8\% of all nodes (yielding at this step $\left|V\right| = 466,610$ nodes). Inspection of removed clusters confirms that these are not relevant to urban issues (medical or chemistry microsimulations for example). Because of the way citation networks are organized and how our dataset was constructed, the full network contains a large part of ``graps'' which consist in several uncited papers citing the same one. We choose thus to work on the ``core'' of the network, which we construct by successively removing nodes of degree less or equal to one until there are no more such nodes.

%%%%%%%%%%%%%%%%%%%%%%
\section*{Mapping the scientific landscape}

@@

We first investigate how network topology varies as a function of horizontal depth.

%%%%%%%%
\begin{figure}
    \centering
    \includegraphics{}
    \caption{}
    \label{fig:modularity-hdepth}
\end{figure}
%%%%%%%%

We then assess that a certain level of overlap exists between the sub-networks originating from different initial requests. As shown in Fig.~\ref{fig:coverage-hdepth}

%%%%%%%%
\begin{figure}
    \centering
    \includegraphics{figures/coveragesubnws.png}
    \caption{}
    \label{fig:coverage-hdepth}
\end{figure}
%%%%%%%%

\subsection*{Reconstructing endogenous disciplines}

% cluster sizes
The distribution of community size is fat-tailed in a power-law manner
